---
title: "TidyTemplate"
date: 2020-09-23
output: html_output
editor_options: 
  chunk_output_type: console
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(tidymodels)
theme_set(theme_minimal())

```

## Load and Explore Data

```{r Load}
## Download and explore data

tt <- tt_load("2020-09-22")

climbers <- tt$members


climbers %>%
  group_by(peak_name) %>%
  summarise(death_rate = mean(died == TRUE),
            total_climbs = n()) %>%
  arrange(desc(death_rate)) %>%
  filter(death_rate < .8) %>%
  ggplot(aes(x = total_climbs, y = death_rate)) +
  geom_point() +
  ggrepel::geom_text_repel(aes(label = if_else(death_rate > .05 | total_climbs > 500, peak_name, ""))) +
  scale_x_log10() +
  scale_y_continuous(label = scales::percent_format()) +
  labs(y = "Death Rate", x = "Number of Climbers", title = "Mortality Rate of Himalayan Mountains")
```


```{r Load}
climbers %>%
  group_by(expedition_role) %>%
  summarise(death_rate = mean(died == TRUE),
            total_climbs = n()) %>%
  arrange(desc(death_rate)) %>%
  filter(death_rate < .8) %>%
  ggplot(aes(x = total_climbs, y = death_rate)) +
  geom_point() +
  ggrepel::geom_text_repel(aes(label = if_else(death_rate > .1 | total_climbs > 100, expedition_role, ""))) +
  scale_x_log10() +
  scale_y_continuous(label = scales::percent_format()) +
  labs(y = "Death Rate", x = "Number of Climbers", title = "Mortality Rate by Expedition Role")
```


```{r Load}
climbers %>%
  group_by(citizenship) %>%
  summarise(death_rate = mean(died == TRUE),
            total_climbs = n()) %>%
  arrange(desc(death_rate)) %>%
  filter(death_rate < .8) %>%
  ggplot(aes(x = total_climbs, y = death_rate)) +
  geom_point() +
  ggrepel::geom_text_repel(aes(label = if_else(death_rate > .025 | total_climbs >1000, citizenship, ""))) +
  scale_x_log10() +
  scale_y_continuous(label = scales::percent_format()) +
  labs(y = "Death Rate", x = "Number of Climbers", title = "Mortality Rate by Citizenship")
```


```{r Load}
climbers %>%
  filter(season != "Unknown",
         year >= 1950) %>%
  mutate(decade = 10 * (year %/% 10)) %>%
  group_by(decade,season) %>%
  summarise(death_rate = mean(died == TRUE),
            total_climbs = n()) %>%
  filter(death_rate < .2) %>%
  ggplot(aes(x = decade, y = death_rate)) +
  geom_point(aes(size = total_climbs)) +
  geom_line() +
  facet_wrap(~season) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "Decade", y = "Death Rate", size = "Total Climbers", title = "Mortality Rate of Himalayan Mountain Climbers Over Time")


```


## Preprocessing for Modeling

```{r}
library(tidymodels)


climbers_model <- climbers %>%
  select(age,peak_name, year, season, sex, citizenship, expedition_role, hired, solo, oxygen_used, died) %>%
  mutate(died = if_else(died == TRUE, "died","survived")) %>%
  mutate(across(where(is.logical),as.numeric)) %>%
  mutate(across(where(is.character),as.factor)) %>%
  filter(!is.na(peak_name),
         !is.na(citizenship),
         !is.na(sex),
         !is.na(expedition_role))


map(climbers_model, ~sum(is.na(.)))


set.seed(1988)

climbers_split = rsample::initial_split(climbers_model, strata = died)
climbers_train <- rsample::training(climbers_split)
climbers_test <- rsample::testing(climbers_split)

climber_folds <- rsample::vfold_cv(data = climbers_train, strata = died)
```

GLMNET Model

```{r}
library(themis)

usemodels::use_glmnet(died ~ ., data = climbers_train)


glmnet_recipe <-
  recipe(formula = died ~ ., data = climbers_train) %>%
  step_medianimpute(age) %>%
  step_other(peak_name, citizenship, expedition_role) %>%
  step_novel(all_nominal(), -all_outcomes()) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors(), -all_nominal()) %>%
  step_upsample(died)


glmnet_spec <-
  logistic_reg(penalty = tune(), mixture = tune()) %>%
  set_mode("classification") %>%
  set_engine("glmnet")

glmnet_workflow <-
  workflow() %>%
  add_recipe(glmnet_recipe) %>%
  add_model(glmnet_spec)

glmnet_grid <- tidyr::crossing(penalty = 10^seq(-6, -1, length.out = 20),
                               mixture = c(0.05, 0.2, 0.4, 0.6, 0.8, 1))




library(doParallel)
registerDoParallel()


glmnet_tune <-
  tune_grid(glmnet_workflow,
            resamples = climber_folds,
            grid = glmnet_grid,
            control = control_grid(save_pred = TRUE))



autoplot(glmnet_tune)

best_glmnet <-  select_best(glmnet_tune, metric = "roc_auc")

collect_predictions(glmnet_tune) %>%
  inner_join(best_glmnet, by = c("penalty","mixture"))

glmnet_pred <- collect_predictions(glmnet_tune) %>%
  inner_join(best_glmnet, by = c("penalty","mixture"))

glmnet_pred %>%
  group_by(id) %>%
  roc_curve(died, .pred_died) %>%
  autoplot()


glmnet_pred %>%
  conf_mat(died, .pred_class)




```


Random Forest model

```{r}


usemodels::use_ranger(died ~., data = climbers_train)


ranger_recipe <-
  recipe(formula = died ~ ., data = climbers_train) %>%
  step_medianimpute(age) %>%
  step_other(peak_name, citizenship, expedition_role) %>%
  step_upsample(died)



ranger_spec <-
  rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>%
  set_mode("classification") %>%
  set_engine("ranger")

ranger_workflow <-
  workflow() %>%
  add_recipe(ranger_recipe) %>%
  add_model(ranger_spec)


ranger_grid <- grid_max_entropy(min_n(), finalize(mtry(), climbers_train), size = 6)

set.seed(60402)



ranger_tune <-
  tune_grid(ranger_workflow,
            resamples = climber_folds,
            grid = ranger_grid,
            control = control_grid(save_pred = TRUE))



ranger_tune %>%
  collect_metrics() %>%
  arrange(desc(mean))

autoplot(ranger_tune)

## Get best model and perform resampling

best_ranger <-  select_best(ranger_tune, metric = "roc_auc")

ranger_pred <- collect_predictions(ranger_tune) %>%
  inner_join(best_ranger, by = c("min_n", "mtry"))

ranger_pred %>%
  group_by(id) %>%
  roc_curve(died, .pred_died) %>%
  autoplot()

ranger_pred %>%
  conf_mat(died, .pred_class)



```


earth model

```{r}

usemodels::use_earth(died ~ ., data = climbers_train)

earth_recipe <-
  recipe(formula = died ~ ., data = climbers_train) %>%
  step_medianimpute(age) %>%
  step_other(peak_name, citizenship, expedition_role) %>%
  step_novel(all_nominal(), -all_outcomes()) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_zv(all_predictors()) %>%
  step_smote(died)


earth_spec <-
  mars(num_terms = tune(), prod_degree = tune(), prune_method = "none") %>%
  set_mode("classification") %>%
  set_engine("earth")

earth_workflow <-
  workflow() %>%
  add_recipe(earth_recipe) %>%
  add_model(earth_spec)

earth_grid <- tidyr::crossing(num_terms = 2 * (1:6),
                              prod_degree = 1:2)



earth_tune <-
  tune_grid(earth_workflow,
            resamples = climber_folds,
            grid = earth_grid,
            control = control_grid(save_pred = T))



collect_metrics(earth_tune) %>%
  arrange(desc(mean))

autoplot(earth_tune)

best_earth <-  select_best(earth_tune, metric = "roc_auc")

earth_pred <- collect_predictions(earth_tune) %>%
  inner_join(best_earth, by = c("num_terms", "prod_degree"))

earth_pred %>%
  group_by(id) %>%
  roc_curve(died, .pred_died) %>%
  autoplot()

earth_pred %>%
  conf_mat(died, .pred_class)

```

```{r}

usemodels::use_xgboost(died ~ ., data = climbers_train)

xgboost_recipe <-
  recipe(formula = died ~ ., data = climbers_train) %>%
  step_medianimpute(age) %>%
  step_other(peak_name, citizenship, expedition_role) %>%
  step_novel(all_nominal(), -all_outcomes()) %>%
  step_dummy(all_nominal(), -all_outcomes(), one_hot = TRUE) %>%
  step_zv(all_predictors()) %>%
  themis::step_upsample(died)



xgboost_spec <-
  boost_tree(trees = 1000, min_n = tune(), tree_depth = tune(), mtry = tune(),
             learn_rate = tune(), loss_reduction = tune(), sample_size = tune()) %>%
  set_mode("classification") %>%
  set_engine("xgboost")

xgboost_workflow <-
  workflow() %>%
  add_recipe(xgboost_recipe) %>%
  add_model(xgboost_spec)

xgboost_grid <-  grid_max_entropy( min_n(), tree_depth(), finalize(mtry(),climbers_train),
             learn_rate(), loss_reduction(), sample_size = sample_prop(), size = 20)


set.seed(36923)

library(doParallel)
registerDoParallel()


xgboost_tune <-
  tune_grid(xgboost_workflow,
            resamples = climber_folds,
            grid = xgboost_grid,
            control = control_grid(save_pred = T))



collect_metrics(xgboost_tune) %>%
  arrange(desc(mean))

autoplot(xgboost_tune)

best_xgboost <-  select_best(xgboost_tune, metric = "roc_auc")

xgboost_pred <- collect_predictions(xgboost_tune) %>%
  inner_join(best_xgboost, by = c("mtry",
                                  "min_n",
                                  "tree_depth",
                                  "learn_rate",
                                  "loss_reduction",
                                  "sample_size"))

xgboost_pred %>%
  group_by(id) %>%
  roc_curve(died, .pred_died) %>%
  autoplot()

xgboost_pred %>%
  conf_mat(died, .pred_class)

save.image(file = "climbers_tidymodels.RData")

```